Synthesizing Quotient Lenses

### 2018-03-08 papers/QuotientLenses/main.tex
pass through sections 5 and 6

(* allow any permutation of fields interspersed with arbitrary whitespace *)
let bib_fields = perm (label, bib_author, bib_title) with (project (wsp+ . "," . wsp+) $\to$ ",\n")
let bibtex = "@book{" . bib_fields . "}"
\end{lstlisting}
\caption{QRE specification of \bibtex{} records. }
\label{fig:example-qre}
\end{figure}
\subsection{QRE Lenses and QRE Lens Synthesis}
At this point we already have a tool for synthesizing bijective lenses from a
pair of regular expressions and a set of example input-output pairs, as well as
the language of QREs for defining equivalence relations on regular expressions.
An obvious next step in simplifying the quotient lens programming is thus to use
the bijective lens synthesis procedure as a subroutine for a quotient lens
synthesis procedure. This new procedure should take two QREs and a set
of example input-output pairs as input, compute the canonical data formats on
both sides of the lens from the pair of QREs, map the example input-output pairs
to their canonical representatives to get a set of examples in canonical form,
and then invoke the bijective lens synthesis procedure on the canonical data
formats and the new set of examples that are in canonical form.

@@

\fi
\section{Synthesizing QRE Lenses}
\label{synth}
We next demonstrate how to use previous results from~\cite{optician} to
synthesize a quotient lens 
$q: c \Leftrightarrow c_1$ from the QREs $c, c_1$ and a set of input-output pairs
$\{(x_1, y_1),\allowbreak \ldots,\allowbreak (x_n, y_n)\}$ where the $x_i$'s are in $W(c)$ and the
$y_i$'s are in $W(c_1)$. We wish $q$ to map the equivalence class of $x_i$ to the
equivalence class of $y_i$ and vice versa:
\begin{align*}
q.get(x_i) &\equiv_{c_1} y_i \text{, and}\\
q.put(y_i) &\equiv_c x_i
\end{align*}

\bcp{Another question about choice of variables: if we rename $c$ and $c_1$
  to $P$ and $Q$, could we also choose names like $p$ and $q$ for their
  elements?  (I guess $q$ was used to range over QRE-lenses, so we can't use
  that suggestion without further renaming...)}

Our approach in synthesizing QRE lenses is guided by Theorem~\ref{normal
form}, which says that, if there is a derivation $q :
c \Leftrightarrow c_1$ of a QRE lens, then there exists a bijective lens $\ell :
K(c) \Leftrightarrow K(c_1)$ such that
\begin{align*}
\llbracket q \rrbracket.get &= \llbracket \ell \rrbracket\circ \canonizer(c)\\
\llbracket q \rrbracket.put &= \llbracket \ell \rrbracket^{-1} \circ
\canonizer(c_1).
\end{align*}

For our synthesis task, since the $x_i$'s are in $W(c)$ and the $y_i$'s are
in $W(c_1)$, then ${x_i}' = \canonizer(c)(x_i)$ is in $K(c)$ and ${y_i}' =
\canonizer(c_1)(y_i)$ is in $K(c_1)$. Consequently in order to synthesize the
desired quotient lens $q: c \Leftrightarrow c_1$ that is consistent with the
input-output examples $\{(x_1, y_1), \ldots, (x_n, y_n)\}$ it suffices to
synthesize a bijective lens $l :
K(c) \Leftrightarrow K(c_1)$ that is consistent with the canonized examples
$\{({x_1}', {y_1}'), \ldots, ({x_n}', {y_n}')\}$

Optician~\cite{optician} showed that given regular expressions
$R, S$ and a set of input-output pairs $\{(r_1, s_1),\allowbreak \ldots,\allowbreak (r_n, s_n)\}$, if
there is a derivation $\ell : R \Leftrightarrow S$ of a bijective lens that
maps $r_i$ to $s_i$ in the forward direction and $s_i$ to $r_i$ in the backward
direction, then there is an algorithm that computes a bijective lens $\ell' :
R' \Leftrightarrow S'$ such that $R \equiv R'$, $S \equiv S'$, $\ell'$ maps
$r_i$ to $s_i$ in the forward direction and $s_i$ to $r_i$ in the backward
direction, and $\llbracket \ell \rrbracket = \llbracket \ell'
\rrbracket$.

This implies that if there is a derivation $q : c \Leftrightarrow
c_1$ of a QRE lens as well as a derivation $\ell : K(c) \Leftrightarrow K(c_1)$
that is consistent with the canonized examples $\{({x_1}', {y_1}'),\allowbreak
\ldots,\allowbreak ({x_n}', {y_n}')\}$, then we can synthesize a quotient lens $q$ of
type $c \Leftrightarrow c_1$ that is consistent with $\{({x_1}', {y_1}'),
\ldots, ({x_n}', {y_n}')\}$. More concretely, in order to synthesize a
quotient lens $q: c \Leftrightarrow c_1$, that is consistent with the examples
$\{(x_1, y_1),\allowbreak \ldots, (x_n, y_n)\}$ it suffices to perform the following
procedure:
\begin{enumerate}
  \item
  Compute $K(c)$, $K(c_1)$,
  \item
  Compute $\{({x_1}', {y_1}'), \ldots, ({x_n}', {y_n}')\}$ where ${x_i}' =
  \canonizer(c)(x_i)$ and ${y_i}' = \canonizer(c_1)(y_i)$,
  \item
  Synthesize a bijective lens $\ell : K(c) \Leftrightarrow K(c_1)$ that is
  consistent with the examples $\{({x_1}', {y_1}'),\allowbreak
  \ldots,\allowbreak ({x_n}',
  {y_n}')\}$
  \item
  Return $q = \kw{rquot}(\kw{lquot}(c, \ell), c_1) : c \Leftrightarrow
  c_1$
\end{enumerate}
For example, recall that in the \bibtex{} to EndNote transformation we had the
QREs \cd{bibtex} and \cd{endnote} describing a \bibtex{} record and an Endnote
record respectively. We also had the example pair \cd{(bib_example ,
end_example)} where

\begin{lstlisting}
bib_example =
"@Book {Doe,
Author = \"John Doe\",
Title = {Generic Title},
}"
\end{lstlisting}

\noindent and

\begin{lstlisting}
end_example =
"%0 Book
%T Generic Title
%A John Doe
%F Doe"
\end{lstlisting}

Since the transformation between the kernel of \cd{bibtex} and the kernel of
\cd{endnote} that maps \cd{bib_example} to \cd{end_example} can be
expressed by a bijective lens, we can synthesize a lens that maps between the
QREs \cd{bibtex} and \cd{endnote}. We do this using the expression

@@

@@

We have integrated QREs and quotient lens synthesis into Boomerang. With this
integration, QREs and quotient lens synthesis tasks can be expressed alongside
existing Boomerang terms.

Our evaluations serve to answer the following questions:
\begin{enumerate}
  \item Does synthesizing quotient lenses from QREs permit an easier development
  process than existing approaches?

@@

\subsection{Benchmark Suite Construction}
Our benchmarks were constructed from two sources:
\begin{enumerate}
  \item 39 of our benchmarks were derived from the Optician system.  These
  benchmarks were a combination of custom benchmarks, benchmarks derived from
  FlashFill, and benchmarks derived from Augeas.  Many of the formats mapped
  between had to be modified to work with the bijectivity constraints
  Optician required.  Commonly, one format had whitespace where the other did
  not allow for whitespace.  To address the bijectivity issues surrounding this,
  we added whitespace to the format that typically did not have whitespace.  We
  were able to remove these alterations in adopting these benchmarks to our
  system.

  \item 8 of our benchmarks were derived from data.gov.  Data.gov contains
  many transformations between formats.  In addition to the formats previously
  adopted for Optician, we synthesized transformers between these formats.
\end{enumerate}

@@

\begin{figure}[t]
\includegraphics{generated-graphs/asts.eps}
\caption{Count of benchmark programs definable using a given AST count. We find
that it takes far fewer AST nodes to define benchmark lenses using QRE
synthesis than with Optician without QRE synthesis or without synthesis. }
\label{fig:asts}
\end{figure}

To test if our tool eases the development process, we compare the following
number of AST nodes.
%
\begin{itemize}
  \item[\QRESize{}] is the number of AST nodes in the QRE
  specifications involved in QRE lens synthesis task.
  \item[\CanonizerAndSpecSize{}] is the number of
  AST nodes in $W(q)$ summed with the number of AST nodes in $\canonizer(q)$,
  for each QRE used in QRE lens synthesis task.
  \item[\LensAndSpecSize{}] is the number of AST nodes
  in $W(q)$ summed with the number of AST nodes in the synthesized QRE lens.
\end{itemize}

The value of \QRESize{} represents the difficulty in writing our benchmarks.

The value of \CanonizerAndSpecSize{} represents the difficulty in writing our
benchmarks within the Optician system. This is only an approximation, as both
$W(q)$ and $\canonizer(q)$ are automatically generated from the QRE, instead of
written by a human.

The value of \LensAndSpecSize{} represents the difficulty in writing our
benchmarks in Boomerang with neither QREs nor synthesis. This is also an
approximation, as $W(q)$ and $\canonizer(q)$ and the bijective lens are all
automatically generated.

Figure~\ref{fig:asts} shows the number of benchmarks written with at most a
given number of AST nodes. For a given benchmark, we found \QRESize{} to be
about half as large as \CanonizerAndSpecSize, and about a third as large as
\LensAndSpecSize{}. This demonstrates that introducing QREs saves programmers
significant effort compared to both Optician and basic Boomerang.

@@

Optician was only able to synthesize fully bijective data transformations.
Synthesizing from QREs loosens that restriction, making it possible to
synthesize data which is bijective when put in canonical form. To determine the
increase of expressivity, we analyzed:
\begin{enumerate}
  \item How many data formats in Optician's benchmark suite required modifications
  to be bijective.
  \item How many data formats in Optician's benchmark suite required modifications
  to be bijective when put in canonical form.
\end{enumerate}

We found that 10 of Opticians 39 benchmarks required modifications to be
bijective. We found that 4 of Optician's 39 benchmarks required modifications to
be in bijective correspondence when in canonical form. Furthermore, we found
that all of the data.gov examples required no modifications to be in bijective
correspondence when in canonical form. While we did not put the data.gov
examples into a form suitable for Optician, we found only 4 of the 8 data.gov
examples did not require canonization, so making suitable for Optician would
have required a modification of 4 of the data.gov formats.

@@

To be used in a development environment, we found it important to validate
performance did not significantly degrade from Optician. To validate this, we
tested the runtimes when:
\begin{itemize}
  \item[\OpticianRuntime{}] Optician runs on its own benchmarks.
  \item[\SystemOnOptician{}] Our system runs on Optician's benchmarks.
  \item[\SystemOnBenchmarks{}] Our system runs on its own benchmarks.
\end{itemize}

We summarize the results of these experiments in Figure~\ref{fig:times}.
Our system was
able to synthesize all of Optician's benchmarks at a speed competitive with
Optician.  There was additional overhead in calculating $W(q)$ and $K(q)$,
resulting in a very slight decrease in performance.

Also notable was one of the data.gov examples, which converted demographic
statistics by zip code from xml form into rdf form. This merely took longer
because it was an exceptionally large format, converting over 40 fields. We
added in additional fields to make this format suitable for Optician, which
took over 15 seconds synthesize the modified lens.

>>>>

(* allow any permutation of fields interspersed with arbitrary whitespace *)
let bib_fields = perm (label, bib_author, bib_title) with (project (wsp+ . "," . wsp+) $\to$ ",\n")
let bibtex = "@book{" . bib_fields . "}"
\end{lstlisting}
\caption{QRE specification of \bibtex{} records. }
\label{fig:example-qre}
\end{figure}
\subsection{QRE Lenses and QRE Lens Synthesis}
\label{sec:examplesynth}
At this point we already have a tool for synthesizing bijective lenses from a
pair of regular expressions and a set of example input-output pairs, as well as
the language of QREs for defining equivalence relations on regular expressions.
An obvious next step in simplifying the quotient lens programming is thus to use
the bijective lens synthesis procedure as a subroutine for a quotient lens
synthesis procedure. This new procedure should take two QREs and a set
of example input-output pairs as input, compute the canonical data formats on
both sides of the lens from the pair of QREs, map the example input-output pairs
to their canonical representatives to get a set of examples in canonical form,
and then invoke the bijective lens synthesis procedure on the canonical data
formats and the new set of examples that are in canonical form.

@@

\fi
\section{Synthesizing QRE Lenses}
\label{synth}

QRE lenses address some of the limitations of bijective lenses because a single
lens program expresses both the canonizers and the transformation between kernel
languages simultaneously, which reduces programmer effort.  But we can go even
further by recognizing that the type structure of QRE lenses contains
information that can be exploited to automatically synthesize lenses from their
types.  Rather than writing the QRE lens manually, the programmer can instead
specify the desired behavior of a lens by giving its interface types and providing
examples, if necessary, to disambiguate among possible implementations.  This
way of constructing lenses can often be much simpler than building them by
hand---see the example of lens synthesis in
Section~\ref{sec:examplesynth}.

The Optician framework~\cite{optician} showed how to do such lens synthesis in
the case for bijective lenses. Here we show how to reduce QRE lens synthesis to
that case, so that we can re-use the Optician algorithm but in the more
expressive context of QRE lenses.  The basic idea is straightforward: we run the
Optician algorithm to synthesize a lens between the kernels of two QREs and then
apply the canonizers at the edges to recover a lens between the whole
languages.  This simple strategy turns out to be remarkably effective, and the
idea of using Optician in this way is part the impetus for the design of our QRE
lenses.

The Optician synthesis algorithm works as follows. Given regular expressions $R$
and $S$ and a set of example input--output pairs
$\{(r_1, s_1),\allowbreak \ldots,\allowbreak (r_n, s_n)\}$, Optician will try to
find a bijective lens $\ell : R \Leftrightarrow S$ that that agrees on the
examples, \textit{i.e.}, it maps $r_i$ to $s_i$ in the forward direction and
$s_i$ to $r_i$ in the backward direction.  The Optician algorithm is guaranteed
to succeed if such a lens exists.    \saz{We should probably mention what happens if the examples aren't sufficient to uniquely
  determine the lens.}

% equivalent regular expressions $R' \equiv R$ and $S' \equiv S$

% if there is a , $\ell'$
% derivation 
% then there is an algorithm that computes a bijective lens
% $\ell' : R' \Leftrightarrow S'$ such that 
% maps $r_i$ to $s_i$ in the forward direction and $s_i$ to $r_i$ in the backward
% direction, and $\llbracket \ell \rrbracket = \llbracket \ell' \rrbracket$.

In our setting, we want to synthesize a quotient lens $q: c \Leftrightarrow c_1$
from the QREs $c_1$ and $c_2$ and a set of example input-output pairs
$\{(x_1, y_1),\allowbreak \ldots,\allowbreak (x_n, y_n)\}$ where the $x_i$'s are
in $W(c_1)$ and the $y_i$'s are in $W(c_2)$. We furthermore wish $q$ to map the
equivalence class of $x_i$ to the equivalence class of $y_i$ and vice versa:
\begin{align*}
q.get(x_i) &\equiv_{c_2} y_i \text{, and}\\
q.put(y_i) &\equiv_{c_1} x_i
\end{align*}

\bcp{Another question about choice of variables: if we rename $c$ and $c_1$
  to $P$ and $Q$, could we also choose names like $p$ and $q$ for their
  elements?  (I guess $q$ was used to range over QRE-lenses, so we can't use
  that suggestion without further renaming...)}\saz{I like the idea of doing a
  consistent renaming of the variables in the paper following this suggestion.
  I would also suggest not using so many primes. }

Our approach to synthesizing QRE lenses is guided by Theorem~\ref{normal form},
which says that, if there is a derivation $q : c_1 \Leftrightarrow c_2$ of a QRE
lens, then there exists a bijective lens $\ell : K(c_1) \Leftrightarrow K(c_2)$
such that
\begin{align*}
\llbracket q \rrbracket.get &= \llbracket \ell \rrbracket\circ \canonizer(c_1)\\
\llbracket q \rrbracket.put &= \llbracket \ell \rrbracket^{-1} \circ
\canonizer(c_2).
\end{align*}

For the QRE examples, the $x_i$'s are in $W(c_1)$ and the $y_i$'s are
in $W(c_2)$, so we can construct ${x'_i} = \canonizer(c_1)(x_i)$ in $K(c_1)$ and ${y'_i} =
\canonizer(c_2)(y_i)$ is in $K(c_2)$. To synthesize the
desired quotient lens $q: c_1 \Leftrightarrow c_2$ that is consistent with the
input-output examples $\{(x_1, y_1), \ldots, (x_n, y_n)\}$ it suffices to
synthesize a bijective lens $l :
K(c_1) \Leftrightarrow K(c_2)$ that is consistent with the canonized examples
$\{({x'_1}, {y'_1}), \ldots, ({x'_n}, {y'_n})\}$ and then apply the canonizers
at the outside.  QRE lens synthesis is

% This implies that if there is a derivation $q : c \Leftrightarrow
% c_1$ of a QRE lens as well as a derivation $\ell : K(c) \Leftrightarrow K(c_1)$
% that is consistent with the canonized examples $\{({x_1}', {y_1}'),\allowbreak
% \ldots,\allowbreak ({x_n}', {y_n}')\}$, then we can synthesize a quotient lens $q$ of
% type $c \Leftrightarrow c_1$ that is consistent with $\{({x_1}', {y_1}'),
% \ldots, ({x_n}', {y_n}')\}$. More concretely, in order to synthesize a
% quotient lens $q: c \Leftrightarrow c_1$, that is consistent with the examples
% $\{(x_1, y_1),\allowbreak \ldots, (x_n, y_n)\}$ it suffices to perform the following
% procedure:
\begin{enumerate}
  \item
  Compute $K(c_1)$, $K(c_2)$,
  \item
  Compute $\{({x'_1}, {y'_1}), \ldots, ({x'_n}, {y'_n})\}$ where ${x'_i} =
  \canonizer(c_1)(x_i)$ and ${y'_i} = \canonizer(c_2)(y_i)$,
\item Synthesize, using Optician's algorithm, a bijective lens
  $\ell : K(c_1) \Leftrightarrow K(c_2)$ that is consistent with the examples
  $\{({x'_1}, {y'_1}),\allowbreak \ldots,\allowbreak ({x'_n}, {y'_n})\}$
  \item
  Return $q = \kw{rquot}(\kw{lquot}(c_1, \ell), c_2) : c_1 \Leftrightarrow
  c_2$
\end{enumerate}

\saz{Do we want to state formally the completeness of this algorithm?}

Returning to the \bibtex{} to EndNote transformation of
Section~\ref{sec:examplesynth}, we had the QREs \cd{bibtex} and \cd{endnote}
describing a \bibtex{} record and an Endnote record respectively. We also had
the example pair \cd{(bib_example , end_example)} where

\begin{minipage}{3in}
\begin{lstlisting}
bib_example =
"@Book {Doe,
 Author = \"John Doe\",
 Title = {Generic Title},
}"
\end{lstlisting}
\end{minipage}
\begin{minipage}{3in}
\begin{lstlisting}
end_example =
"%0 Book
 %T Generic Title
 %A John Doe
 %F Doe"
\end{lstlisting}
\end{minipage}

Since the transformation between the kernel of \cd{bibtex} and the kernel of
\cd{endnote} that maps \cd{bib_example} to \cd{end_example} can be
expressed as a bijective lens, we can synthesize a lens that maps between the
QREs \cd{bibtex} and \cd{endnote}, using the code below:

@@

In this case, one example is sufficient to uniquely determine the desired
transformation.\saz{Is that true?}

\saz{We might want to say something more precise about the running time or
  expected performance of this algorithm, perhaps to set up the evaluation
  section.  Something along the lines of: ``Although the algorithm is complete
  in theory, in the worst case, it may take an unbounded [???IS THIS CORRECT?]
  amount time to search for a satisfactory lens.  Nevertheless, this algorithm
  performs well in practice on numerous examples, as we show in the next
  section.}

@@

We have implemented QREs and the quotient lens synthesis algorithm described
above as part of Boomerang. Thanks to this integration, QREs and quotient lens
synthesis tasks can now be expressed alongside existing Boomerang terms, and we
can combine the two kinds of lenses seamlessly in the same application.
Intuition suggests that the increased expressiveness of QREs should make many
lens programming tasks easier, but we would like more quantitative understanding
of that claim.  We thus conducted several experiments that aim to provide
answers to the following questions:

\begin{enumerate}
  \item Does synthesizing quotient lenses from QREs permit an easier development
  process than existing approaches?

@@

\subsection{Benchmark Suite Construction}

To evaluate our QRE implementation, we constructed a set of lens synthesis tasks
drawn from two sources:
\begin{enumerate}
\item 39 of our benchmarks were adapted from the evaluation of the original
  Optician system.  These benchmarks are a combination of custom benchmarks,
  benchmarks derived from FlashFill~\cite{}\saz{TODO:cite}, and benchmarks
  derived from Augeas~\cite{}\saz{TODO:cite}.  In the pre-QRE version of
  Optician, many of these data formats had to be modified to work with the
  bijectivity constraints that Optician required.  For instance, when one format
  permits whitespace where the other does not allow it, the original versions of
  benchmark formats were modified to allow more whitespace, thereby restoring
  bijectivity.  With the new QRE support, we were able to remove these
  alterations when porting these benchmarks.  (This experience alone suggests
  that QREs make the lens development process easier.)

\item The remaining eight benchmarks were derived from
  data.gov.\footnote{\url{http://data.gov}} Data.gov contains many
  transformations between formats.  In addition to the formats previously
  used for Optician, we synthesized transformers between these formats.\saz{This
    description is not specific enough.  What are the data formats for?  How did
    we pick which ones to try?}
\end{enumerate}

@@

\begin{figure}[t]
\includegraphics{generated-graphs/asts.eps}
\caption{Count of benchmark programs definable using a given AST count. We find
that it takes far fewer AST nodes to define benchmark lenses using QRE
synthesis than with Optician without QRE synthesis or without synthesis.\saz{Why
is the bar for ``All QRE Benchmarks'' at 14? Shouldn't it be at 47 = 39+8? Are
we not testing all QRE benchmarks?}}
\label{fig:asts}
\end{figure}

To get a sense of how much the new expressiveness that combining QRE lenses with
synthesis saves in terms of programmer effort, we compared the amount of code
that would need to be written for each of the lens benchmarks.  We use number of
nodes in the abstract syntax tree as a measure of code size, and calculated
three quantities for each lens synthesis benchmark task.  \saz{Say something
  about why there are only 14 benchmarks?}
%
\begin{itemize}
  \item[\QRESize{}] the number of AST nodes in the QRE
  specifications (including examples)\saz{CHECK THIS}
  \item[\CanonizerAndSpecSize{}]  the number of
  AST nodes in $W(q)$ summed with the number of AST nodes in $\canonizer(q)$
  \item[\LensAndSpecSize{}] the number of AST nodes
  in $W(q)$ summed with the number of AST nodes in the synthesized QRE lens.
\end{itemize}

The value of \QRESize{} is our measure of the programmer burden for
writing our benchmarks---it measures the number of AST nodes the programmer has
to write using the full expressiveness of QREs and synthesis.

The value of \CanonizerAndSpecSize{} estimates the burden for writing our
benchmarks within the Optician system, without QREs but still using
synthesis. This is only an approximation, as both $W(q)$ and $\canonizer(q)$ are
automatically generated from the QRE, and it is possible that a human-written
version might be smaller.

The value of \LensAndSpecSize{} estimates the burden in writing our benchmarks
in Boomerang, using neither QREs nor synthesis. This value is also an
approximation, as $W(q)$ and $\canonizer(q)$ and the bijective lens are all
automatically generated.

Figure~\ref{fig:asts} shows the number of benchmarks that can be expressed using
at most a given number of AST nodes. For a given benchmark, we found \QRESize{}
to be about half as large as \CanonizerAndSpecSize, and about a third as large
as \LensAndSpecSize{}. This demonstrates that introducing QREs saves programmers
significant effort compared to both Optician and basic Boomerang.

@@

The pre-QRE version of Optician was only able to synthesize fully bijective data
transformations.  Synthesizing from QREs loosens that restriction, making it
possible to synthesize data which is bijective when put in canonical form. To
determine the increase of expressivity, we analyzed:
\begin{enumerate}
  \item How many data formats in Optician's benchmark suite required modifications
  to be bijective.
  \item How many data formats in Optician's benchmark suite required modifications
  to be bijective when put in canonical form.
\end{enumerate}

We found that 10 of Optician's 39 benchmarks required modifications to be
bijective. We found that 4 of Optician's 39 benchmarks required modifications to
be in bijective correspondence when in canonical form.\saz{Is this somehow where
  the 14 = 10+4 points in Fig. 10 come from?}\saz{Do these 10 and 4 intersect?}
Furthermore, we found that all eight of the data.gov examples required no
modifications to be in bijective correspondence when in canonical form. While we
did not put the data.gov examples into a form suitable for pre-QRE Optician, we
found that only four of the eight data.gov examples did not require
canonization, so making them suitable for pre-QRE Optician would have required 
changes to of four of the data.gov formats.  Here again, we see that QREs add
significant expressive power.

@@

To assess the performance impact of adding QRE support to Optician, we measured
the running time of the following configurations:
\begin{itemize}
  \item[\OpticianRuntime{}] pre-QRE Optician running on its own benchmarks.
  \item[\SystemOnOptician{}] Optician with QREs running on the pre-QRE benchmarks.
  \item[\SystemOnBenchmarks{}] Optician with QREs running on its own benchmarks.
\end{itemize}

We summarize the results of these experiments in Figure~\ref{fig:times}.  The
new version of Optician was able to synthesize all of pre-QRE Optician's
benchmarks at a speed competitive with the old version.  There is a small amount
of additional overhead introduced by QREs when calculating $W(q)$ and $K(q)$,
resulting in a very slight decrease in performance.

Also notable is one of the data.gov examples, which converts demographic
statistics by zip code from xml form into rdf form. This took longer because it
was an exceptionally large format, converting over 40 fields.\saz{We need to
  explain more detail here.  I guess this is about the last benchmark in Fig. 11
  (b), which takes a long time?}  We added in additional fields to make this
format suitable for Optician, which took over 15 seconds synthesize the modified
lens.\saz{This needs more context, explanation.}
